---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r setoptions, echo = TRUE}

```

## Loading and preprocessing the data

```{r}
library(tidyverse)
library(dplyr)
library(tibble)
library(tidyr)
library(readr)

if(!file.exists('activity.csv')){
    unzip('activity.zip')
}

data <- read_csv("activity.csv")

names(data)
str(data)

```

## What is mean total number of steps taken per day?

```{r}
library(ggplot2)
library(dplyr)
data2 <- data%>%
  na.omit()%>%
  group_by(date)%>%
  summarize(sum_step = sum(steps , na.rm = TRUE))

qplot(data2$sum_step, geom="histogram" ,  binwidth=500 , 
      xlab = "Daily steps" ,
      main = "Histogram of the total number of steps taken each day") 

# Calculate and print mean of total steps
mean_total_steps <- mean(data2$sum_step , na.rm = TRUE)
print(mean_total_steps)


# Calculate and print median of total steps
median_total_steps <- median(data2$sum_step,  na.rm = TRUE)
print(median_total_steps)
```

- Mean:  10766.19

- Median: 10765

## What is the average daily activity pattern?

```{r}
library(ggplot2)
library(dplyr)

data3 <- data%>%
  group_by(interval)%>%
  summarize(sum_interval = sum(steps , na.rm = TRUE))

ggplot(data3 , aes(x = interval , y = sum_interval)) + 
  geom_line()+
  labs(x= "Interval" , y = "Average steps" ,  
       title = "Average daily activity pattern" )

data3%>%
  arrange(desc(sum_interval))

```
The "835" 5-minute interval contains the maximum number of steps

## Imputing missing values

To fill in the missing values, I decided to use the mean of 5 minute interval.
```{r}
library(ggplot2)
library(dplyr)

#1. Number of missing values
na_values <- sum(is.na(data$steps))

#2-3. Fill in  missing values

neosteps <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))
dat <- data%>% 
   group_by(interval)%>%
   mutate(steps = neosteps(steps))

dat <- dat%>%
  group_by(date)%>%
  summarize(sum_step_na = sum(steps))

qplot(dat$sum_step_na, geom="histogram" ,  binwidth=500 , 
      xlab = "Daily steps" ,
      main = "Histogram of the total number of steps taken each day with Na values fixed") 

# Calculate and print mean of total steps with NA values fixed
mean_total_steps_na <- mean(dat$sum_step_na)
print(mean_total_steps_na)


# Calculate and print median of total steps with Na values fixed 
median_total_steps_na <- dat
  median(dat$sum_step_na)

```


- Number of missing values: 2304
- Mean numbers of daily steps with missing values: 10766.19
- Median numbers of daily steps with missing values: 10766.19

So, the mean is the same, while median are increased a bit.

## Are there differences in activity patterns between weekdays and weekends?

```{r}
library(ggplot2)
library(dplyr)
days <- weekdays(data$date)
dt <- data%>%
    mutate(days = weekdays(date))%>%
    mutate(daytype =  ifelse(days == "Monday" | days == "Tuesday" | days == "Wednseday" | days == "Thursday" | days == "Friday", "Weekdays" , "Weekend"))

f_data <- dt%>%
  group_by(interval)%>%
  mutate( avg =  mean(steps, na.rm = TRUE) )

ggplot(f_data , aes(x = interval , y = avg , color = daytype)) +
  geom_line()+
  facet_grid(~daytype)

```


